Clinical Deployment Environments: The Five Pillars of Translational Machine Learning for Health

### 2022-07-02 main.tex
corrected caption and diagram

\begin{figure}[h!]
\begin{center}
\includegraphics[width=10cm]{assets/emap.png}% This is a *.eps file
\end{center}
\caption{Our real-world development is performed on the  Experimental Medicine Application Platform (EMAP). EMAP is a clinical laboratory within which ML4H researchers can iteratively build, test and gather feedback from the bedside. It unifies the data and the tools for off-line and online development of ML4H models. In brief, EMAP builds a patient orientated SQL database from Health Level 7 version 2 (HL7v2) messages that are being exchanged between hospital systems. HL7v2 messages are ubiquitous in health care, and the \textit{de facto} standard for internal communication.\newline Rather than multiple pairwise connections between different hospital electronic systems, an integration engine acts as a single hub that routes HL7 messages, and where necessary translates to ensure compatibility. EMAP copies each message passing through the integration engine to a PostgreSQL database, the \textit{Immutable Data Store (IDS)} (see Figure 1 and the \textbf{(numbers)} in the following sentences that refer to objects in the figure).\newline A \textit{message reader} \textbf{(2)} processes each live message to an interchange format so that downstream processing is insulated from local HL7 implementation. Separately, the \textit{table reader} \textbf{(5)} processes historical data (e.g. from the reporting database) to the same interchange format. Live messages take priority over historical messages in a queue that feeds the \textit{event processor} \textbf{(3)}. This links each message to a patient and a hospital visit, makes appropriate updates for out of order messages, and merges when separate identifiers are recognised to represent the same patient. A full audit trail is maintained. Each event updates a second live PostgreSQL database, the \textit{User Data Store} (UDS) \textbf{(4)}. The hospital hosts Jupyter and RStudio servers, and a Linux development environment is provided that allows docker deployment, installation of analysis libraries and frameworks, exposes SSH and HTTPS services, and allows user verification against the hospital active directory. \textbf{(5)} A typical workflow might include investigation and experimentation in a Jupyter Notebook with data from the UDS, then using a small network of docker containers to run the development script, log outputs to a testing database, and report to users via email or a locally hosted web application or dashboard.  A fuller explanation is available in the Electronic Supplementary Material (Section 2: EMAP data flows).}\label{fig:1}
\end{figure}

>>>>

\begin{figure}[h!]
\begin{center}
\includegraphics[width=10cm]{assets/emap.png}% This is a *.eps file
\end{center}
\caption{Our real-world development is performed on the  Experimental Medicine Application Platform (EMAP). EMAP is a clinical laboratory within which ML4H researchers can iteratively build, test and gather feedback from the bedside. It unifies the data and the tools for off-line and online development of ML4H models (see Figure 1 and the \textbf{(numbers)} in the following sentences that refer to objects in the figure).\newline
In brief, EMAP builds a patient orientated SQL database from Health Level 7 version 2 (HL7v2) messages that are being exchanged between hospital systems. HL7v2 messages are ubiquitous in health care, and the \textit{de facto} standard for internal communication. Rather than multiple pairwise connections between different hospital electronic systems, an integration engine acts as a single hub that routes HL7 messages, and where necessary translates to ensure compatibility.\newline
EMAP copies each message passing through the integration engine to a PostgreSQL database, the \textit{Immutable Data Store (IDS)} \textbf{(1)}. A \textit{message reader} \textbf{(2)} processes each live message to an interchange format so that downstream processing is insulated from local HL7 implementation. Separately, the \textit{table reader} \textbf{(6)} processes historical data (e.g. from the reporting database) to the same interchange format. Live messages take priority over historical messages in a queue that feeds the \textit{event processor} \textbf{(3)}. This links each message to a patient and a hospital visit, makes appropriate updates for out of order messages, and merges when separate identifiers are recognised to represent the same patient. A full audit trail is maintained. Each event updates a second live PostgreSQL database, the \textit{User Data Store} (UDS) \textbf{(4)}. The hospital hosts Jupyter and RStudio servers, and a Linux development environment is provided that allows docker deployment, installation of analysis libraries and frameworks, exposes SSH and HTTPS services, and allows user verification against the hospital active directory. \textbf{(5)} A typical workflow might include investigation and experimentation in a Jupyter Notebook with data from the UDS, then using a small network of docker containers to run the development script, log outputs to a testing database, and report to users via email or a locally hosted web application or dashboard.  A fuller explanation is available in the Electronic Supplementary Material (Section 2: EMAP data flows).}\label{fig:1}
\end{figure}
